---
title: "STATISTICS 4-6"
author: "Luckson"
date: "16/09/2019"
output: html_document
---

## Exercise 4.1
```{r}
vec1 <- c(6,9,7,3,6,7,9,6,3,6,6,7,1,9,1) 
vec1
vec1==6
vec1 >= 6
vec1<6+2
vec1 !=6
#part B
vec2 <-c (3,6,7,9,6,3,6,6,7,1,9,1)
vec2
newvec <- array(data=c(vec2),dim =c(2,2,3))
newvec
#part C
newvec <= (6/2)+4
newvec <- c(newvec)+2
#terd <= 6
## D part 
elem <- diag(x=10)
elem
elem == 0

newvec == T
## E part 
elem2 <- diag(elem)
elem2
elem2 != T

```
## Exercise 4.2
```{r}
# A answer
foo <- c(7,1,7,10,5,9,10,3,10,8)
foo
elem <- foo > 5| foo == 2
elem
# B answer
barvec <- c(8,8,4,4,5,1,5,6,6,8)
barvec
barvec2 <- bar<=6 & bar != 4
barvec2
# C answer
elem1 <- foo & barvec2
elem1
# D answer
bazvec <- c(foor + barr)
bazvec
bazelem <- c(baz <= 14) | c(baz !=15)
bazelem
vecelem <- ((baz/foo) > 4 | (baz/foo) <= 2)
vecelem
```
## Exercise 4.3
```{r}
foovec1 <- c(7,5,6,1,2,10,8,3,8,2)
foovec1
# A answer
bar <- foovec1[foovec1>= 5]
bar
# ii part
foovec2 <-  foovec1[(foovec1< 5)]
foovec2
# B answer
baz <- matrix(data=c(bar),nrow=2,ncol=3)
baz
# i part
sqval<- (baz[1,2])^2
sqval

baz <- c(baz)
baz
baz [c(5,6)] <- c(sqval,sqval)
baz
# ii part
baz[(baz<=25)&(baz > 4)]
baz
# C answer
vec1 <- c(10,5,1,4,7,4,3,3,1,3,4,3,1,7,8,3,7,3)
vec1
qux <- array(vec1,dim=c(3,2,3))
qux
# i of C
which((x=qux ==3)| (x=qux ==4),arr.ind =T)
qux
# ii of C
qux<- qux[(qux <3) ||(qux >= 7)] <- 100 
qux
# D answer

foo <- c(7,1,7,10,5,9,10,3,10,8)
foo
ext <- foo[c(F,T)]
ext <- foo[c(0,1)]
why|why not

```
## Exercise 4.4
```{r}
# A answer
paste("The quick brown fox\n\tjumped over\n\t\tthe lazy dogs")
# B answer
num1 <- 4
num1
num2 <- 0.75
num2
paste("The result of multipying ",num1,"by",num2,"is",num1*num2)
# C answer
mac1<- c("/Users","/tdavies","/Documents","/RBook/")
mac1
paste(mac1[1],mac1[3],mac1[4],"lzvirikuzhe",sep="")
mac2 <- c("/lzvirikuzhe")
mac2
paste (mac1[1],mac1[3],mac2[1],mac1[4],sep="")
# D answer
bar <- c("How","much","wood","could","a","woodchuck","chuck")
bar
length(x=bar)
bar2 <- c("if","a","woodchuck","could","chunk","wood")
bar2
newbar <- paste(bar2[1],bar2[2],bar2[3],bar2[4],bar2[5],bar[6],bar[1],bar[2],bar[3],bar[4],bar[5],bar[6],bar[7])
newbar
gsub(pattern="wood",replacement="metal",x=newbar)
# E answer
foo <- c("Two 6-packs for $12.99!")
foo
qual <- paste(foo[1],foo[2],foo[3],foo[4])
qual
substr(x=foo,start=5,stop=10)

gsub(pattern="12.99",replacement="10.99",x=qual)
```
## Exercise 4.5
```{r}
# A answer
sex.character <- c("female","male","male","male","female","female","female","male","male","male","male","female","male","female","female","female","male","male","male","male")
sex.character
sex.character.fac <- factor(x=sex.character)
sex.character.fac
party.name  <- c("National","Labour","Greens","Maori","Other")
party.name
party.name.fac  <- factor(x=party.name)
party.name.fac
levels  <- levels(x=party.name.fac) <- c("1","2","3","4","5")
# B answer
sex.num <- c(name)
sex.num
sex.num.fac <- factor(x=sex.num)
sex.num.fac
levels(x=sex.num.fac)
levels(x=sex.num.fac) <- c("Female","Male")
sex.num.fac

party.num <- c(name)
party.num
party.num.fac <- factor(x=party.num)
party.num.fac
levels(x=party.num.fac)
levels(x=party.num.fac) <- c("Labour","Greens","Other","National","Maori")
party.num.fac
sex.num.fac

party.char.fac <- factor(x=party.num.fac, ordered = TRUE)
party.char.fac
# Since individuals have been grouped by the party of their choice without any order so making no sense in ordering.
## R seems to have put sequence based on the individual listing given at the start.
# C answer

party.males <- (party.num.fac[sex.num.fac=="Male"])  ## returns factor vector of parties for males only
party.males
sex.national <- (sex.num.fac[party.num.fac== "National"])
sex.national

party.num.fac ## after adding 6 more individuals, their party's and gender type
sex.num.fac ## after adding 6 more individuals, their party's and gender type

conflevel <- c(93, 55, 29, 100,52, 84, 56, 0, 33, 52, 35, 53, 55, 46, 40, 40, 56, 45, 64, 31, 10, 29, 40,
95, 18, 61)
conflevel
confbreak <- c(0,30,70,100) ## (0 to less 30 -low),(30 to less than 70- Moderate),(70 to equal to 100 -high)
confbreak
cut (x=conflevel,breaks = confbreak)
cut (x=conflevel,breaks = confbreak,right = F)
cut (x=conflevel,breaks = confbreak,right = F,include.lowest = T)
confname <- c("Low","Moderate","High")
confname
confidencelevels <- cut (x=conflevel,breaks = confbreak,right = F,include.lowest = T,labels = confname)
confidencelevels

conflevellabor <- (confidencelevels[party.num.fac== "Labour"])
conflevellabor

conflevelNational <- (confidencelevels[party.num.fac== "National"])
conflevelNational

#labour confidence is between Moderate-High, and Nationals confidence is Low to Moderate ##

```
# Exercise 5.1
```{r}
# A answer
fig<- seq(from=-4,to=4,length= 20)
fig
veclog<-matrix(c(F,T,T,T,F,T,T,F,F),nrow=3, ncol=3)
veclog
vecchar<-c("don", "quixote")
vecchar
vecfac<- factor(x=c("LOW","MED","LOW","MED","MED","HIGH"))
vecfac
L   <- list(fig,veclog,vecchar,vecfac)
L

# i elements extraction nrow(2:1), ncol(2:3)
L[[2]][2:1,2:3]

#ii Sub to overwrite quixote and don

L[[3]][1]<-sub(pattern="don", replacement="Don",x=L[[3]][1])
L[[3]][2]<-sub(pattern="quixote", replacement="Quixote",x=L[[3]][2])
L[[3]]
cat("\"Windmills! ATTACK!\"\n"," ","", "-\\" ,L[[3]],"/-")

#iii numbers > than 1
valg <- L[[1]][L[[1]]>1]
valg
#iv using which factor vector assigned to MED
which(x=L[[4]]=="MED")

# B answer

newlist <- list(facs=L[[4]],figs=c(3,2.1,3.3,4,1.5,4.9),oldlist=L[1:3])
newlist
#i Extract elements of facs > = 3
Egreater <- newlist@facs[3][newlist@figs>= 3]
Egreater
#ii add new member
newlist@flags <- rep(x=newlist@oldlist[[2]][,3], times=2)
newlist@flags
#iii extraction for entries of figures =false
entries<- newlist@nums[][!newlist@flags]
entries
# Overwrite oldlist
newlist@oldlist[[3]]<- "Don Quixote"
```
# Exercise 5.2

```{r}
# A answer
framed <- data.frame(person=c("Stan","Francine","Steve","Roger","Hayley","Klaus"),
         sex=factor(x=c("M","F","M","M","F","M")),   
         funny=factor(x=c("High","Med","Low","High","Med","Med")
                                   levels=c("Low","Med","High")),stringsAsFactors=FALSE)
framed

# B answer 
framed@age <- c(41,41,15,1600,21,60)
framed

# C answer
framed <- framed [c(1,4,2,3)]
framed
# D answer
mydata <- data.frame(person=c("Peter","Lois","Meg","Chris","Stewie","Brian")
age=c(42,40,17,14,1,7),sex=factor(x=c("M","F","F","M","M", "M")),funny=factor(x=c("High","High","Low","Med","High","Med")) age.mon=c(504,480,204,168,12,84),stringsAsFactors=FALSE)
mydata
mydata2 <- mydata[,-5] 
mydata2

# E answer
mydataframe <- rbind(mydata2,framed)
mydataframe
# F answer
mydataframe[mydataframe@sex=="F"&(mydataframe@funny=="Med"|mydataframe@funny=="High"),c("person","age")]

# G answer
mydataframe[substr(x=mydataframe@person,start=1,stop=1)=="S",]

# Exercise 6
foo <- c(13563,-14156,-14319,16981,12921,11979,9568,8833,-12968,8133)
foo[which(x=is.infinite(foo^75))]
foo[which(x=foo^75 != -Inf)]
bar <-matrix(c(77875.40, 27551.45, 23764.,30,-36478.88,-35466.25, -73333.85, 36599.69 -70585.69,-39803.81, 55976.34, 76694.82,47032.00),nrow=3,ncol=4)
which(x=is.nan(bar^65/Inf))
bar[is.nan(bar^67+Inf)==FALSE]
bar[bar^67!=-Inf]
bar[bar==-Inf | is.finite(bar^67)]
```

# Exercise 6.2

```{r}
foo <- c(4.3,2.2,NULL,2.4,NaN,3.3,3.1,NULL,3.4,NA)
length(foo)
which(x=is.na(x=foo))
is.null(x=foo)
is.na(x=foo[8])+4/NULL
mylist <- list(c(7,7,NA,3,NA,1,1,5,NA))
names(mylist) <-'alpha'
exists('mylist$beta')
b <- list(which(x=is.na(mylist$alpha)))
c(mylist,beta=b)
```


# Exercise 6.3

```{r}
class(foo <- array(data=1:36,dim=c(3,3,4)))
class(bar <- as.vector(foo))
class(baz <- as.character(bar))
class(qux <- as.factor(baz))
class(quux <- bar+c(-0.1,0.1))
a=is.numeric(foo)+is.integer(foo)
b=is.numeric(bar)+is.integer(bar)
c=is.numeric(baz)+is.integer(baz)
d=is.numeric(qux)+is.integer(qux)
e=is.numeric(quux)+is.integer(quux)
factor(c(a,b,c,d,e))
as.numeric(c(a,b,c,d,e))
m <- matrix(seq(from=2,to=13),nrow=3,ncol=4)
as.character(m)
m2 <- matrix(c(34,23,33,42,41,0,1,1,0,0,1,2,1,1,2),nrow=5,ncol=3)
dfm2 <- as.data.frame(m2)
dfm2
as.logical(dfm2[,2]) 
as.factor(dfm2[,3])





